---
title: "Test"
author: "Jihoon Kim"
date: "11/3/2017"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R  

R is 'GNU S', a freely available language and environment for statistical computing and graphics which provides a wide variety of statistical and graphical techniques

- linear and nonlinear modeling
- statistial tests
- time series analysis
- classification
- clustering
- genomics


Download and install R (precompiled binary distribution) from the nearest R mirror site, <https://cran.stat.ucla.edu>




## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```


## Look and feel of R 

```{r}
1+2
2*2
2*2*2*2
2^4
```

## More calculations in R

```{r}
log(16, 2)
pi
pi*(3^2)
sin(pi/2)
```


## Simple plot in R

```{r}
x = c(-10:10)
y = x^3
plot(x, y)
```


## Make the figure fancier by adding more plotting parameters

```{r}
x = c(-10:10)
y = x^3
plot(x, y, pch=1, lty=1, type="b", col="blue")
```


## Try changing parameter values 

```{r}
x = c(-5:10)
y = x^3
plot(x, y, pch=15, lty=2, type="b", col="orange")
```




## R Studio 

RStudio is a set of integrated tools designed to help you be more prudictive with R. It includes 

- console
- syntax-highlighting editor that supports direct code execution
- plotting
- history
- debugging 
- workspace management. 

Download and install RStudio from <https://www.rstudio.com/products/rstudio/download>


## R Studio text editor
R Studio -> File -> New File -> R Script to create an R code file, type commands below in the text file, click the first line, and click Run button three times.

```{r, eval=FALSE}
x = c(-10:10)
y = x^3
plot(x, y, pch=1, lty=1, type="b", col="blue")
```


## R Objects
Text goes here




## R Packages

R packages are collections of functions, data, and complied code in a ready-to-use format with enriched documentation. 

Use R Studio to install R packages 

- package name 1
- package name 2
- package name 3


## Survival analysis







## Principal Component Analysis (PCA)


## R code for PCA with automobile data


## PCA plot with automobile data

## Application of PCA to GWAS


## R code for Population Stratification using PCA


## Plot of population stratification 


## Population structure within Europe (Novembre et al. 2008)